The role of habitat quality in host metapopulations exposed to environmentally persistent pathogens

### 2015-06-09 Manuscript/manuscript.tex
Clarity edits

We assume that patch population size is proportional to patch quality ($A_i$), and that disease-induced mortality reduces the size of infectious populations by a factor of $\nu$.  Building on the framework developed by \cite{Hanski2003} then, the rates at which patch \emph{i}, with quality $A_i$, is colonized by susceptible and infectious individuals, respectively, is given by its connectivity:
\begin{align}
C_{Si} &= A_i^{\xi_{im}} \sum_{i\neq j }\phi(S_j)A_j^{\xi_{em}}\exp(-D d_{ij}),\\
C_{Ii} &= A_i^{\xi_{im}} \sum_{i\neq j }\phi(I_j)(\nu A_j) ^{\xi_{em}}\exp(-D d_{ij}),
\label{connectivity}
\end{align}
where $\phi(X_j)$ is an indicator function that is $1$ if patch $j$ is in state $X$ and is $0$ otherwise; $\xi_{im}$ and $\xi_{em}$ control how rates of immigration and emigration, respectively, scale with patch quality; $D$ is the inverse of mean dispersal distance; and $d_{ij}$ is the distance between patches $i$ and $j$.  Essentially, $C_{Xi}$ sums the colonization effort (or propagule pressure) from all patches in state $X$ to focal patch $i$.

@@

To explore the relative influence of low and high quality habitat on host and pathogen dynamics, we simulated the spread of a range of pathogens in metapopulations with either high or low quality habitat distributions (with patch qualities ranging from 0.76 to 1.77, and from 0.23 to 1.24, respectively).  Specifically, we implemented a full-factorial design in which disease-induced mortality ($\nu$) and probability of direct infection ($\delta$) were each varied over 10 values (ranging from 0.1 to 1, and 0 to 0.9, respectively).  In addition, pathogen longevity was varied over 3 values (ranging from 0.3 to 3 times the expected residence time of a healthy population on a unit quality patch, i.e., 10 units of time).

>>>>

We assume that the host population size supported by a patch is proportional to its quality ($A_i$), and that disease-induced mortality reduces the size of infectious populations by a factor of $\nu$.  Building on the framework developed by \cite{Hanski2003} then, the rates at which patch \emph{i}, with quality $A_i$, is colonized by susceptible and infectious individuals, respectively, is given by its connectivity:
\begin{align}
C_{Si} &= A_i^{\xi_{im}} \sum_{i\neq j }\phi(S_j)A_j^{\xi_{em}}\exp(-D d_{ij}),\\
C_{Ii} &= A_i^{\xi_{im}} \sum_{i\neq j }\phi(I_j)(\nu A_j) ^{\xi_{em}}\exp(-D d_{ij}),
\label{connectivity}
\end{align}
where $\phi(X_j)$ is an indicator function that is $1$ if patch $j$ is in state $X$ and is $0$ otherwise; $\xi_{im}$ and $\xi_{em}$ control how rates of immigration and emigration, respectively, scale with patch quality; $D$ is the inverse of mean dispersal distance; and $d_{ij}$ is the distance between patches $i$ and $j$.  Essentially, $C_{Xi}$ sums the colonization effort (or propagule pressure) from all patches in state $X$ to focal patch $i$.

@@

To explore the relative influence of low and high quality habitat on host and pathogen dynamics, we simulated the spread of a range of pathogens in metapopulations with patch quality distributions shifted towards either high or low quality habitat.
These two distributions were chosen to have equal variance and an overlapping core of patches around unit quality to roughly simulate alternative management actions that favor expansion to include either additional high or low quality habitat (with qualities ranging from 0.76 to 1.77, and from 0.23 to 1.24 respectively).  
We then implemented a full-factorial design on each quality distribution in which the pathogen's disease-induced mortality ($\nu$) and probability of direct infection ($\delta$) were each varied over 10 values (ranging from 0.1 to 1, and 0 to 0.9, respectively).  In addition, pathogen longevity was varied over 3 values (ranging from 0.3 to 3 times the expected residence time of a healthy population on a unit quality patch, i.e., 10 units of time).
